---
title: "SummitLake"
author: "Leah Walker"
date: "16/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(plyr)
library(dplyr)
library(xml2)
library(stringr)
library(tidyr)
library(ggplot2)
library(Rmisc)
library(lme4)
library(lmerTest)
library(rsortie)

source("./R/clean_summit_lake_data.R")
source("../Carbon/CarbonFunctions.R") # Need to have the Carbon repo downloaded onto your system
source("../Carbon/AllometryFunctions.R") # same
source("./R/SL_field_calc_functions.R")

BA.function <- function(DBH) {pi*(DBH/200)^2}

path_outputs <- "./Outputs/"

```
-----------------------------------------
I didn't have time to create nice functions for each of the prep steps (sorry!), so I've just included them all here, in the correct order. You can skip down to line 426 for where the carbon calculations and analysis begin. -LW
-----------------------------------------

Clean data
*Read in the Summit Lake field data here, if running other code processing the data*
```{r}
################## Read in the data ######################
#rm(list=ls())

### Run these three likes of code if running any other code to do with the field data
summit.lk.dat <- read.csv("./Inputs/Data/SummitLakeData.csv")

summit.lk.dat <- summit.lk.dat %>% 
  mutate(DBH_09_live = as.numeric(DBH_09_live), 
         DBH_19_live = as.numeric(DBH_19_live))

###

dbhClSize <- 2
MinDBHClass <- 2 # 2.0 - 3.9 is the first DBH class
MaxDBHClass <- ceiling(max(na.omit(summit.lk.dat$DBH_c_92_live)))
PlotArea <- 0.05 # in ha

# Clean the data to be in a format readable by SORTIE
dat <- clean_summit_lk_dat(dbhClSize = dbhClSize, MinDBHClass = MinDBHClass,
                           MaxDBHClass = MaxDBHClass, PlotArea = PlotArea,
                           raw_data = summit.lk.dat)

```



Prepare Inputs
*Don't need to run unless making new parameter files*
```{r}
sl_fl <- "./Inputs/FileLists/"

sl_basexmls <- "./Inputs/ParameterFiles/BaseFiles/"
sl_newvals <- "./Inputs/ParameterValues/"
sl_newxmls <- "./Inputs/ParameterFiles/"

sl_variable_names <- treelistDfn(initname = "Init.Dens_", numDigits = 1, diamMin = 0, diamMax = 88, diamInc = 2) 

### SBS-01 ###

lstFiles_01 <- (paste0(sl_fl, "InitSummitLake_01.csv"))

makeFiles(lstFiles = lstFiles_01, path_basexmls = sl_basexmls, path_newvals = sl_newvals, path_newxmls = sl_newxmls, variable_names = sl_variable_names)

### SBS-0506 ###

lstFiles_0506 <- (paste0(sl_fl, "InitSummitLake_0506.csv"))

makeFiles(lstFiles = lstFiles_0506, path_basexmls = sl_basexmls, path_newvals = sl_newvals, path_newxmls = sl_newxmls, variable_names = sl_variable_names)

```



Run SORTIE
*Don't need to run unless wanting to re-run the simulations*
```{r} 

### SBS-01 ###
xmls2run <- grep("SBS-01e-", list.files("./Inputs/ParameterFiles/"), value = TRUE)

for (ix in 1:length(xmls2run)) {
  runSortie(paste0(sl_newxmls, xmls2run[ix]),0)
}



### SBS-0506 ###
xmls2run2 <- grep("SBS-0506f-", list.files("./Inputs/ParameterFiles/"), value = TRUE)

for (ix in 1:length(xmls2run2)) {
  runSortie(paste0(sl_newxmls, xmls2run2[ix]),0)
}

```



Extract and read outputs in R (not using - not sure if working properly)
*Alana may have fixed!* - Sept 2022
```{r}
#path_outputs <- "./Outputs/"

#dt <- data.table()
#yrs <- seq(0,1)
#units <- unique(summit.lk.dat$Plot)

#for(i in 1:length(units)){
#  units_i <- units[i]
#  for(ii in 1:length(yrs)){
#  dt_unit <- fread(paste0(path_outputs, "Extracted/Ext_SBS-01e-summit", units_i, "-p_det_", yrs[ii]))
#  dt_unit[, ':='(timestep = yrs[ii], unit = units[i])]
#  dt <- rbind(dt, dt_unit)
#  }
#  write.csv(dt, paste0(path_outputs,"csv/summit",units_i,"_out.csv"), row.names = FALSE)
#}

```



Extract outputs in SORTIE and read in R

*Start here to extract the outputs and make a csv file with all the output data. If you already have the csv's on your system, you can read them in here*

*Treatments*
Control: 3, 6, 9, 10, 16, 17, 20
Medium: 4, 8, 11, 12, 15, 18 (medium residual BA)
Low: 5, 7, 13, 14, 19, 24 (low residual BA)

```{r}

##################################################################################################################
### If you have already written the CSV for theses files you can skip to lines 198 (SBS-01) and 251 (SBS-0506) ###
##################################################################################################################

u92 <- c(3,5,6,7,8,9,10,11,12,13,14,16,17,18,19,20,24)
u94 <- c(4,15)
yrs92 <- seq(0,27)
yrs94 <- seq(0,25)

treatment <- data.table(unit = c(3, 6, 9, 10, 16, 17, 20, 4, 8, 11, 12, 15, 18, 5, 7, 13, 14, 19, 24),
                        treatment = c(rep("ctrl", times = 7), rep("med", times = 6), rep("low", times = 6)))


### SBS-01 ###


files_92_01 <- paste0("Extracted/ext_SBS-01e-summit",u92,"-p_det_")
files_94_01 <- paste0("Extracted/ext_SBS-01e-summit",u94,"-p_det_")

out_01 <- data.table()

for(i in 1:length(u92)){
  units_i <- u92[i]
  for(ii in 1:length(yrs92)){
    dt <- fread(paste0(path_outputs, files_92_01[i], yrs92[ii]), sep = "\t", 
                header = TRUE, na.strings = "--", skip = 1)
    dt[,':='(timestep = yrs92[ii], unit = units_i)]
    dt <- dt[X >50 & X <150 & Y >50 & Y <150]
    out_01 <- rbindlist(list(out_01, dt), use.names = TRUE, fill = TRUE)
  }
}

for(i in 1:length(u94)){
  units_i <- u94[i]
  for(ii in 1:length(yrs94)){
    dt <- fread(paste0(path_outputs, files_94_01[i], yrs94[ii]), sep = "\t", 
                header = TRUE, na.strings = "--", skip = 1)
    dt[,':='(timestep = yrs94[ii], unit = units_i)]
    dt <- dt[X >50 & X <150 & Y>50 & Y <150]
    out_01 <- rbindlist(list(out_01, dt), use.names = TRUE, fill = TRUE)
  }
}

out_01 <- merge(out_01, treatment, by = "unit")

#write.csv(out_01, paste0(path_outputs, "csv/SummitLake_output_01.csv"), row.names = FALSE)

####################
### SKIP TO HERE ###
####################

out_01 <- read.csv("./Outputs/csv/SummitLake_output_01.csv")
out_01 <- as.data.table(out_01)

# Data collection for Unit 4 and 15 started in 1994, not 1992 - DO NOT run these two lines if testing initiation
out_01$timestep[out_01$unit == 4] <- out_01$timestep[out_01$unit == 4] + 2
out_01$timestep[out_01$unit == 15] <- out_01$timestep[out_01$unit == 15] + 2

# Subset living and dead
out_01_L <- subset(out_01, out_01$Type != "Snag")
out_01_D <- subset(out_01, out_01$Type == "Snag")



### SBS-0506 ###

files_92_0506 <- paste0("Extracted/ext_SBS-0506f-summit",u92,"-p_det_")
files_94_0506 <- paste0("Extracted/ext_SBS-0506f-summit",u94,"-p_det_")

out_0506 <- data.table()

for(i in 1:length(u92)){
 units_i <- u92[i]
 for(ii in 1:length(yrs92)){
   dt <- fread(paste0(path_outputs, files_92_0506[i], yrs92[ii]), sep = "\t", 
               header = TRUE, na.strings = "--", skip = 1)
   dt[,':='(timestep = yrs92[ii], unit = units_i)]
   dt <- dt[X >50 & X <150 & Y >50 & Y <150]
   out_0506 <- rbindlist(list(out_0506, dt), use.names = TRUE, fill = TRUE)
 }
}

for(i in 1:length(u94)){
 units_i <- u94[i]
 for(ii in 1:length(yrs94)){
   dt <- fread(paste0(path_outputs, files_94_0506 [i], yrs94[ii]), sep = "\t", 
               header = TRUE, na.strings = "--", skip = 1)
   dt[,':='(timestep = yrs94[ii], unit = units_i)]
   dt <- dt[X >50 & X <150 & Y>50 & Y <150]
   out_0506 <- rbindlist(list(out_0506, dt), use.names = TRUE, fill = TRUE)
 }
}

out_0506 <- merge(out_0506, treatment, by = "unit")

#write.csv(out_0506, paste0(path_outputs, "csv/SummitLake_output_0506.csv"), row.names = FALSE)

####################
### SKIP TO HERE ###
####################

out_0506 <- read.csv("./Outputs/csv/SummitLake_output_0506.csv")
out_0506 <- as.data.table(out_0506)

# Data collection for Unit 4 and 15 started in 1994, not 1992 - DO NOT run these two lines if testing initiation
out_0506$timestep[out_0506$unit == 4] <- out_0506$timestep[out_0506$unit == 4] + 2
out_0506$timestep[out_0506$unit == 15] <- out_0506$timestep[out_0506$unit == 15] + 2

# Subset living and dead
out_0506_L<-subset(out_0506, out_0506$Type != "Snag")
out_0506_D<-subset(out_0506, out_0506$Type == "Snag")

```



Calculate BA per tree and per unit for every timestep - Field and SORTIE - LIVE AND DEAD
*You must run this section if you want to make any plots of live BA*
*Creates a data table of the BA for each unit and timestep*
```{r}
# Field BA LIVE - can skip if csv saved
#SL_calc_field_BA(raw_data = summit.lk.dat, live = TRUE) # You need to run lines 34-37 prior to this line (read in Summit Lake field data)

# Read in LIVE field BA data
field_BA_L <- read.csv("./Outputs/csv/SummitLake_BA_field_L.csv") # Once you've run the SL_calc_field_BA function, you can just read in the data here

# Field BA DEAD - can skip if csv saved
#SL_calc_field_BA(raw_data = summit.lk.dat, live = FALSE) # You need to run lines 34-37 prior to this line (read in Summit Lake field data)

# Read in DEAD field BA data
field_BA_D <- read.csv("./Outputs/csv/SummitLake_BA_field_D.csv")



### SBS-01 ###


# Calculate LIVE BA per tree
out_01_L_BA <- out_01_L[, ':='(BA = BA.function(out_01_L$DBH))]

# Sum BA of live trees per timestep per plot 
out_01_L_BA <- out_01_L_BA %>% 
  group_by(timestep, unit) %>% 
  mutate(BA_unit = sum(BA, na.rm = TRUE))

out_01_L_BA <- out_01_L_BA %>%
  dplyr::select(unit, timestep, treatment, BA_unit)


out_01_L_BA <- as.data.table(out_01_L_BA)
out_01_L_BA <- unique(out_01_L_BA)

write.csv(out_01_L_BA, paste0(path_outputs, "csv/SummitLake_BA_01_L.csv"), row.names = FALSE)


# Calculate DEAD BA per tree
out_01_D_BA <- out_01_D[, ':='(BA = BA.function(out_01_D$DBH))]

# Sum BA of dead trees per timestep per plot 
out_01_D_BA <- out_01_D_BA %>% 
  group_by(timestep, unit) %>% 
  mutate(BA_unit = sum(BA, na.rm = TRUE))

out_01_D_BA <- out_01_D_BA %>%
  dplyr::select(unit, timestep, treatment, BA_unit)

out_01_D_BA <- as.data.table(out_01_D_BA)
out_01_D_BA <- unique(out_01_D_BA)

write.csv(out_01_D_BA, paste0(path_outputs, "csv/SummitLake_BA_01_D.csv"), row.names = FALSE)


### SBS-0506 ### 


# Calculate LIVE BA per tree
out_0506_L_BA <- out_0506_L[, ':='(BA = BA.function(out_0506_L$DBH))]

# Sum BA of live trees per timestep per plot 
out_0506_L_BA <- out_0506_L_BA %>% 
  group_by(timestep, unit) %>% 
  mutate(BA_unit = sum(BA, na.rm = TRUE))

out_0506_L_BA <- out_0506_L_BA %>%
  dplyr::select(unit, timestep, treatment, BA_unit)

out_0506_L_BA <- as.data.table(out_0506_L_BA)
out_0506_L_BA <- unique(out_0506_L_BA)

write.csv(out_0506_L_BA, paste0(path_outputs, "csv/SummitLake_BA_0506_L.csv"), row.names = FALSE)


# Calculate DEAD BA per tree
out_0506_D_BA <- out_0506_D[, ':='(BA = BA.function(out_0506_D$DBH))]

# Sum BA of dead trees per timestep per plot 
out_0506_D_BA <- out_0506_D_BA %>% 
  group_by(timestep, unit) %>% 
  mutate(BA_unit = sum(BA, na.rm = TRUE))

out_0506_D_BA <- out_0506_D_BA %>%
  dplyr::select(unit, timestep, treatment, BA_unit)

out_0506_D_BA <- as.data.table(out_0506_D_BA)
out_0506_D_BA <- unique(out_0506_D_BA)

write.csv(out_0506_D_BA, paste0(path_outputs, "csv/SummitLake_BA_0506_D.csv"), row.names = FALSE)


```



Check to make sure SORTIE runs were initiated properly (**VERIFIED, don't need to run**)
```{r}

### Field BA timestep 0
init_field_BA_L <- subset(field_BA_L, timestep == 0) # Read in field data at the top of section 5



### SBS-01 ###

init_01_L_BA <- subset(out_01_L_BA, timestep == 0)

init_01_L_BA <- merge(init_01_L_BA, init_field_BA_L, by = c("unit", "treatment", "timestep"))
names(init_01_L_BA)[names(init_01_L_BA) == "BA_unit.x"] <- "BA_unit"
names(init_01_L_BA)[names(init_01_L_BA) == "BA_unit.y"] <- "field_BA"


# Plot
# Looking for the data points to be on or close to the 1:1 line
png("./Figures/Inital_SBS01_L_BA.png", width=1000, height=1000, units="px",
    pointsize=23, antialias="default")

plot(init_01_L_BA$field_BA, init_01_L_BA$BA_unit, xlim = c(0,33), ylim = c(0,33)) + abline(0,1)

dev.off()



### SBS-0506 ### 

init_0506_L_BA <- subset(out_0506_L_BA, timestep == 0) # Read in field data at the top of section 5

init_0506_L_BA <- merge(init_0506_L_BA, init_field_BA_L, by = c("unit", "treatment", "timestep"))
names(init_0506_L_BA)[names(init_0506_L_BA) == "BA_unit.x"] <- "BA_unit"
names(init_0506_L_BA)[names(init_0506_L_BA) == "BA_unit.y"] <- "field_BA"


# Plot
# Looking for the data points to be on or close to the 1:1 line
png("./Figures/Inital_SBS0506_L_BA1.png", width=1000, height=1000, units="px",
    pointsize=23, antialias="default")

plot(init_0506_L_BA$field_BA, init_0506_L_BA$BA_unit, xlim = c(0,33), ylim = c(0,33)) + abline(0,1)

dev.off()


```


------------------------------------------------------------------
See the BA_plots R script to re-create the basal area (BA) plots

From plotting the final year of the simulation against the field data, we decided to use the SBS-0506 runs over the SBS-01. 

For the rest of this RMD, we only only include the SBS-0506 results. Please see the C_SBS_01 R script for SBS-01 related carbon calculations and plots. 
------------------------------------------------------------------


Read in SORTIE data

```{r}

# Read in the csv file - created in the Data_prep_validation RMD
out_0506 <- read.csv("./Outputs/csv/SummitLake_output_0506.csv")
out_0506 <- as.data.table(out_0506)

# Data collection for Unit 4 and 15 started in 1994, not 1992 - DO NOT run these two lines if testing initiation
out_0506$timestep[out_0506$unit == 4] <- out_0506$timestep[out_0506$unit == 4] + 2
out_0506$timestep[out_0506$unit == 15] <- out_0506$timestep[out_0506$unit == 15] + 2

# Subset living and dead
out_0506_L<-subset(out_0506, out_0506$Type != "Snag")
out_0506_D<-subset(out_0506, out_0506$Type == "Snag")

```




Carbon calculations (LIVE ONLY)
*Read in SORTIE outputs in section 4 before running this code*
The C calculations and plots for the SBS-01 runs can be found in the R script C_SBS_01.R (in the R folder)
```{r}

###############
### FIELD C ###
###############

# Field C LIVE - can skip if csv saved
#SL_calc_field_C(raw_data = summit.lk.dat, live = TRUE) # You need to run lines 46-50 prior to this line (read in Summit Lake field data)

# Read in LIVE field C data
field_C_L <- read.csv("./Outputs/csv/SummitLake_C_field_L.csv") # Once you've run the SL_calc_field_C function, you can just read in the data here



################
### SBS 0506 ###
################


### LIVE ###

###
### You can skip below if the csv files are already created ###
###

# Calculate Carbon per plot per timestep from SORTIE output

out_0506_L_C <- out_0506_L

out_0506_L_C[, Class := 1]

out_0506_L_C$Species[out_0506_L_C$Species == "Subalpine_Fir"] <- "Bl"
out_0506_L_C$Species[out_0506_L_C$Species == "Interior_Spruce"] <- "Sx"
out_0506_L_C$Species[out_0506_L_C$Species == "Paper_Birch"] <- "Ep"
out_0506_L_C$Species[out_0506_L_C$Species == "Western_Larch"] <- "Lw"
out_0506_L_C$Species[out_0506_L_C$Species == "Black_Cottonwood"] <- "Ac"
out_0506_L_C$Species[out_0506_L_C$Species == "Douglas_Fir"] <- "Fd"
out_0506_L_C$Species[out_0506_L_C$Species == "Trembling_Aspen"] <- "At"
out_0506_L_C$Species[out_0506_L_C$Species == "Lodgepole_Pine" ] <- "Pl"


# Create an empty vector
Carbon <- vector()

# Calculate carbon per tree based on species, DBH, height, and tree class in vector
for(i in 1:nrow(out_0506_L_C)){
  Carbon[i] <- TreeCarbonFN(Species = out_0506_L_C[i, Species], DBH = out_0506_L_C[i, DBH], 
                          HT = out_0506_L_C[i, Height], Tree_class = out_0506_L_C[i, Class])
}

# Bind vector to data frame as a column and convert from Kg/tree to Mg/tree
out_0506_L_C[, ':='(C_tree = Carbon/1000)]

# Calculate carbon per unit
out_0506_L_C <- out_0506_L_C %>% 
  group_by(timestep, unit) %>% 
  mutate(C_unit = sum(C_tree, na.rm = TRUE))

#write.csv(out_0506_L_C, paste0(path_outputs, "csv/SummitLake_C_0506_L.csv"), row.names = FALSE)



####################
### SKIP TO HERE ###
####################

out_0506_L_C <- read.csv(paste0(path_outputs, "csv/SummitLake_C_0506_L.csv"))


out_0506_L_C <- out_0506_L_C %>%
  dplyr::select(unit, timestep, C_unit, treatment)

out_0506_L_C <- unique(out_0506_L_C)

out_0506_L_C <- as.data.table(out_0506_L_C)

```



LIVE carbon plots
```{r}

# Compare final year of simulation C to final year of field data C - LIVE ONLY
final_C_0506_L <- subset(out_0506_L_C, timestep == 27)

final_C_0506_L <- as.data.table(final_C_0506_L)


final_field_C_L <- subset(field_C_L, timestep == 27)


final_C_0506_L <- merge(final_C_0506_L, final_field_C_L, by = c("unit", "treatment", "timestep"))
names(final_C_0506_L)[names(final_C_0506_L) == "C_unit.x"] <- "C_unit"
names(final_C_0506_L)[names(final_C_0506_L) == "C_unit.y"] <- "field_C"



lm_C_finalyr0506 <-lm(C_unit ~ field_C - 1, final_C_0506_L)
summary(lm_C_finalyr0506)


# Different colour by treatment
png("./Figures/Final_SBS0506_L_C.png", width=700, height=500, units="px", pointsize=23, antialias="default")

ggplot(final_C_0506_L, aes(x = field_C, y = C_unit, shape = treatment, color = treatment, label = unit)) +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 14), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12)) +
  geom_point(aes(colour = factor(treatment), shape = factor(treatment)), size = 2) +
  geom_text(hjust=-0.5, vjust=-0.5) +
  geom_abline() +
  scale_x_continuous(breaks = seq(30, 120, by = 10), limits = c(30, 120)) +
  scale_y_continuous(breaks = seq(30, 120, by = 10), limits = c(30, 120)) +
  labs(x = "Observed Carbon (Mg/ha)", y = "Predicted Carbon (Mg/ha)", col = "Treatment", shape = "Treatment") +
  scale_shape_manual(values = c(16, 15, 17),
                     breaks = c("ctrl", "med", "low"), 
                     labels = c("Control", "Medium", "Low")) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium", "Low")) +
  geom_abline(slope = coef(lm_C_finalyr0506)[["field_C"]], linetype = 3)

dev.off()



# Carbon over time by treatment

field_C_L_T <- field_C_L

field_C_L_T <- summarySE(field_C_L_T,
                      measurevar="C_unit",
                      groupvars=c("treatment", "timestep"))

field_C_L_TN <- NULL


out_0506_L_C_T <- out_0506_L_C 

out_0506_L_C_T <- summarySE(out_0506_L_C_T,
                      measurevar="C_unit",
                      groupvars=c("treatment", "timestep"))

out_0506_L_C_T$N <- NULL

png("./Figures/C_0506_L_T3.png", width=1600, height=800, units="px",
    pointsize=23, antialias="default")

ggplot(NULL, aes(x = timestep,  y = C_unit, fill = treatment, group = treatment)) + 
  geom_line(data = out_0506_L_C_T, aes(col = treatment), size = 1.2) + 
  theme(axis.line = element_line(colour = "black"), 
        panel.background = element_blank(), 
        legend.position = "top", 
        legend.text = element_text(size = 25),
        text = element_text(size = 28),
        axis.text = element_text(size = 23),
        axis.text.x = element_text(vjust = -1),
        axis.text.y = element_text(hjust = 0.5),
        axis.ticks.length = unit(7, "pt"),
        axis.title.x = element_text(margin = margin(t = 14,  r = 0, b = 1, l = 0), face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  scale_x_continuous(breaks = seq(0, 28, by = 2), expand = expansion(mult = c(0, 0.05))) +
  scale_y_continuous(breaks = seq(0, 120, by = 20), expand = c(0,0), limits = c(0,120)) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium retention", "Low retention")) +
  scale_fill_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium retention", "Low retention")) +
  coord_cartesian(ylim = c(0, 120)) +
  labs(x = "Years post treatment", y = "Carbon (Mg/ha)", col = "Treatment", fill = "Treatment", shape = "Treatment") +
  geom_ribbon(data = out_0506_L_C_T, aes(ymin = C_unit - ci, ymax = C_unit + ci),  alpha=0.1, linetype="dashed",  color="grey") +
  geom_point(data = field_C_L_T, aes(shape = treatment), size = 5 , position = position_dodge(width = 0.5)) +
  scale_shape_manual(values = c(24, 23, 21),
                     breaks = c("ctrl", "med", "low"), 
                     labels = c("Control", "Medium retention", "Low retention")) +
  geom_errorbar(data = field_C_L_T, aes(ymin = C_unit - ci, ymax = C_unit + ci), position = position_dodge(width = 1)) 
  
dev.off()





# Plot all timesteps at once

# Compare final year of simulation C to final year of field data C - LIVE ONLY
all_C_0506_L <- subset(out_0506_L_C, timestep == 0 | timestep ==  2 | timestep == 5 | timestep == 17 | timestep == 27)

all_C_0506_L <- merge(all_C_0506_L, field_C_L, by = c("unit", "timestep", "treatment"))
names(all_C_0506_L)[names(all_C_0506_L) == "C_unit.x"] <- "C_unit_sortie"
names(all_C_0506_L)[names(all_C_0506_L) == "C_unit.y"] <- "C_unit_field"



lm_C_0506_ctrl <-lm(all_C_0506_L[all_C_0506_L$treatment == "ctrl"]$C_unit_sortie ~ all_C_0506_L[all_C_0506_L$treatment == "ctrl"]$C_unit_field - 1)
summary(lm_C_0506_ctrl)

lm_C_0506_med <-lm(all_C_0506_L[all_C_0506_L$treatment == "med"]$C_unit_sortie ~ all_C_0506_L[all_C_0506_L$treatment == "med"]$C_unit_field - 1)
summary(lm_C_0506_med)

lm_C_0506_low <-lm(all_C_0506_L[all_C_0506_L$treatment == "low"]$C_unit_sortie ~ all_C_0506_L[all_C_0506_L$treatment == "low"]$C_unit_field - 1)
summary(lm_C_0506_low)



png("./Figures/All_SBS0506_L_C2.png", width=700, height=500, units="px", pointsize=23, antialias="default")

ggplot(all_C_0506_L, aes(x = C_unit_field, y = C_unit_sortie, shape = treatment, color = treatment)) +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 18), 
        legend.title = element_text(size = 18), 
        legend.text = element_text(size = 17)) +
  geom_point(aes(colour = factor(treatment), shape = factor(treatment)), size = 2) +
  geom_abline() +
  scale_x_continuous(breaks = seq(0, 120, by = 10), limits = c(0, 120)) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), limits = c(0, 120)) +
  labs(x = "Observed Carbon (Mg/ha)", y = "Predicted Carbon (Mg/ha)", col = "Treatment", shape = "Treatment") +
  scale_shape_manual(values = c(16, 15, 17),
                     breaks = c("ctrl", "med", "low"), 
                     labels = c("Control", "Medium retention", "Low retention")) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium retention", "Low retention")) +
  geom_abline(slope = coef(lm_C_0506_ctrl)[["all_C_0506_L[all_C_0506_L$treatment == \"ctrl\"]$C_unit_field"]], colour = "#F8766D") +
  geom_abline(slope = coef(lm_C_0506_med)[["all_C_0506_L[all_C_0506_L$treatment == \"med\"]$C_unit_field"]], colour = "#619CFF") +
  geom_abline(slope = coef(lm_C_0506_low)[["all_C_0506_L[all_C_0506_L$treatment == \"low\"]$C_unit_field"]], colour = "#00BA38")

dev.off()



```


Carbon calculations (DEAD ONLY)
```{r}

###############
### FIELD C ###
###############

# Field C DEAD - can skip if csv saved
SL_calc_field_C(raw_data = summit.lk.dat, live = FALSE) # You need to run lines 34-37 prior to this line (read in Summit Lake field data)

# Read in DEAD field BA data
field_C_D <- read.csv("./Outputs/csv/SummitLake_C_field_D.csv") # Once you've run the SL_calc_field_C function, you can just read in the data here



################
### SBS 0506 ###
################


### DEAD ONLY ### 


# Calculate Carbon per plot per timestep from SORTIE output

out_0506_D_C <- out_0506_D

out_0506_D_C[, Class := 3]

out_0506_D_C$Species[out_0506_D_C$Species == "Subalpine_Fir"] <- "Bl"
out_0506_D_C$Species[out_0506_D_C$Species == "Interior_Spruce"] <- "Sx"
out_0506_D_C$Species[out_0506_D_C$Species == "Paper_Birch"] <- "Ep"
out_0506_D_C$Species[out_0506_D_C$Species == "Western_Darch"] <- "Lw"
out_0506_D_C$Species[out_0506_D_C$Species == "Black_Cottonwood"] <- "Ac"
out_0506_D_C$Species[out_0506_D_C$Species == "Douglas_Fir"] <- "Fd"
out_0506_D_C$Species[out_0506_D_C$Species == "Trembling_Aspen"] <- "At"
out_0506_D_C$Species[out_0506_D_C$Species == "Lodgepole_Pine" ] <- "Pl"

# Create an empty vector
Carbon <- vector()

# Calculate carbon per tree based on species, DBH, height, and tree class in vector
for(i in 1:nrow(out_0506_D_C)){
  Carbon[i] <- TreeCarbonFN(Species = out_0506_D_C[i, Species], DBH = out_0506_D_C[i, DBH], 
                          HT = out_0506_D_C[i, Height], Tree_class = out_0506_D_C[i, Class])
}

# Bind vector to data frame as a column and convert from Kg/tree to Mg/tree
out_0506_D_C[, ':='(C_tree = Carbon/1000)]

# Calculate carbon per unit
out_0506_D_C <- out_0506_D_C %>% 
  group_by(timestep, unit) %>% 
  mutate(C_unit = sum(C_tree, na.rm = TRUE))

#write.csv(out_0506_D_C, paste0(path_outputs, "csv/SummitLake_C_0506_D.csv"), row.names = FALSE)



####################
### SKIP TO HERE ###
####################

out_0506_D_C <- read.csv(paste0(path_outputs, "csv/SummitLake_C_0506_D.csv"))

out_0506_D_C <- out_0506_D_C %>%
  dplyr::select(unit, timestep, C_unit, treatment)

out_0506_D_C <- unique(out_0506_D_C)

out_0506_D_C <- as.data.table(out_0506_D_C)
```



DEAD carbon plots
```{r}
# Compare final year of simulation C to final year of field data C - DEAD ONLY
final_C_0506_D <- subset(out_0506_D_C, timestep == 27)

final_C_0506_D <- as.data.table(final_C_0506_D)

final_field_C_D <- subset(field_C_D, timestep == 27)


final_C_0506_D <- merge(final_C_0506_D, final_field_C_D, by = c("unit", "treatment", "timestep"))
names(final_C_0506_D)[names(final_C_0506_D) == "C_unit.x"] <- "C_unit"
names(final_C_0506_D)[names(final_C_0506_D) == "C_unit.y"] <- "field_C"



lm_C_finalyr0506_D <-lm(C_unit ~ field_C - 1, final_C_0506_D)
summary(lm_C_finalyr0506_D)



# Different colour by treatment
png("./Figures/Final_SBS0506_D_C.png", width=700, height=500, units="px", pointsize=23, antialias="default")

ggplot(final_C_0506_D, aes(x = field_C, y = C_unit, shape = treatment, color = treatment, label = unit)) +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 14), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12)) +
  geom_point(aes(colour = factor(treatment), shape = factor(treatment)), size = 2) +
  geom_text(hjust=-0.5, vjust=-0.5) +
  geom_abline() +
  scale_x_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  labs(x = "Observed Carbon (Mg/ha)", y = "Predicted Carbon (Mg/ha)", col = "Treatment", shape = "Treatment") +
  scale_shape_manual(values = c(16, 15, 17),
                     breaks = c("ctrl", "med", "low"), 
                     labels = c("Control", "Medium", "Low")) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium", "Low")) +
  geom_abline(slope = coef(lm_C_finalyr0506_D)[["field_C"]], linetype = 3)

dev.off()




# Carbon over time by treatment

out_0506_D_C_T <- out_0506_D_C 

out_0506_D_C_T <- summarySE(out_0506_D_C_T,
                      measurevar="C_unit",
                      groupvars=c("treatment", "timestep"))

out_0506_D_C_T$N <- NULL

png("./Figures/C_0506_D_T.png", width=1600, height=800, units="px",
    pointsize=23, antialias="default")

ggplot(NULL, aes(x = timestep,  y = C_unit, fill = treatment, group = treatment)) + 
  geom_line(data = out_0506_D_C_T, aes(col = treatment), size = 1.2) + 
  theme(axis.line = element_line(colour = "black"), panel.background = element_blank(), legend.position = "top")+
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  ylab("Live carbon Mg/ha") + 
  xlab('Years post-treatment') +
  coord_cartesian(ylim = c(0, 30)) +
  geom_ribbon(data = out_0506_D_C_T, aes(ymin = C_unit - ci, ymax = C_unit + ci),  alpha=0.1, linetype="dashed",  color="grey") +
  geom_point(data = summit.lk.dat_C_D_field, aes(shape = treatment), size = 3 , position = position_dodge(width = 0.5)) +
  scale_shape_manual(values = c(24, 21, 23, 22)) +
  geom_errorbar(data = summit.lk.dat_C_D_field, aes(ymin = C_unit - ci, ymax = C_unit + ci), position = position_dodge(width = 1)) 
  
dev.off()


```



Analysis
```{r}

############
### LIVE ###
############

out_0506_L_C <- read.csv(paste0(path_outputs, "csv/SummitLake_C_0506_L.csv"))
field_C_L <- read.csv(paste0(path_outputs, "csv/SummitLake_C_field_L.csv"))



### RSS and RMSE ###

# Compare final year of simulation C to final year of field data C - LIVE ONLY
final_C_0506_L <- subset(out_0506_L_C, timestep == 27)
final_C_0506_L <- as.data.table(final_C_0506_L)

final_field_C_L <- subset(field_C_L, timestep == 27)

final_C_0506_L <- merge(final_C_0506_L, final_field_C_L, by = c("unit", "treatment", "timestep"))
names(final_C_0506_L)[names(final_C_0506_L) == "C_unit.x"] <- "C_unit"
names(final_C_0506_L)[names(final_C_0506_L) == "C_unit.y"] <- "field_C"



rss0506 <- with(final_C_0506_L, sum((C_unit - field_C)^2))
rmse0506 <- with(final_C_0506_L, sqrt(mean((C_unit - field_C)^2)))



### Linear and mixed effects models ###

# Data manipulation
out_0506_L_C_M <- out_0506_L_C
out_0506_L_C_M <- as.data.table(out_0506_L_C_M)

out_0506_L_C_M <- unique(out_0506_L_C_M)


#out_0506_L_C_M[, DataSource := "sortie"]
out_0506_L_C_M <- subset(out_0506_L_C_M, timestep == 0 | timestep ==  2 | timestep == 5 | timestep == 17 | timestep == 27)

field_C_L <- as.data.table(field_C_L)
#summit.lk.dat_C[, DataSource := "field"]




out_0506_L_C_M2 <- merge(out_0506_L_C_M, field_C_L, by = c("unit", "treatment", "timestep"))
names(out_0506_L_C_M2)[names(out_0506_L_C_M2) == "C_unit.x"] <- "sortie"
names(out_0506_L_C_M2)[names(out_0506_L_C_M2) == "C_unit.y"] <- "field"

out_0506_L_C_M2 <- melt(out_0506_L_C_M2, id.vars = c("unit", "treatment", "timestep"), measure.vars = c("sortie", "field"))
names(out_0506_L_C_M2)[names(out_0506_L_C_M2) == "variable"] <- "DataSource"
names(out_0506_L_C_M2)[names(out_0506_L_C_M2) == "value"] <- "C_unit"

#out_0506_L_C_M2 <- rbindlist(list(out_0506_L_C_M, summit.lk.dat_C), use.names = TRUE, fill = TRUE)

out_0506_L_C_M2$unit <- as.character(out_0506_L_C_M2$unit)


# Basic linear models 
datasourcelm <- lm(C_unit ~ DataSource - 1, out_0506_L_C_M2) # forces intercept through zero, tells us the two means
summary(datasourcelm)

datasourcelm2 <- lm(C_unit ~ DataSource, out_0506_L_C_M2) # does not force the intercept through zero, tells us the difference between the two means
summary(datasourcelm2)

t.test(out_0506_L_C_M2[out_0506_L_C_M2$DataSource == "sortie"]$C_unit, out_0506_L_C_M2[out_0506_L_C_M2$DataSource == "field"]$C_unit)
# simple t-test, tells us the two means and if there is a difference between them (p-value)


# both the lm2 and t-test give a p-value of 0.6979 - therefore there is no significant difference between the data from the two data sources


# Linear model for each timestep WITH intercept
sourcetimelm0 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 0]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 0]$DataSource)
summary(sourcetimelm0)

sourcetimelm2 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 2]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 2]$DataSource)
summary(sourcetimelm2)

sourcetimelm5 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 5]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 5]$DataSource)
summary(sourcetimelm5)

sourcetimelm17 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 17]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 17]$DataSource)
summary(sourcetimelm17)

sourcetimelm27 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 27]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 27]$DataSource)
summary(sourcetimelm27)




# Linear model for each timestep WITHOUT intercept
sourcetimelm0 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 0]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 0]$DataSource - 1)
summary(sourcetimelm0)

sourcetimelm2 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 2]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 2]$DataSource - 1)
summary(sourcetimelm2)

sourcetimelm5 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 5]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 5]$DataSource - 1)
summary(sourcetimelm5)

sourcetimelm17 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 17]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 17]$DataSource - 1)
summary(sourcetimelm17)

sourcetimelm27 <- lm(out_0506_L_C_M2[out_0506_L_C_M2$timestep == 27]$C_unit ~ out_0506_L_C_M2[out_0506_L_C_M2$timestep == 27]$DataSource - 1)
summary(sourcetimelm27)


### MIXED EFFECTS MODELS ###

allfactors <- lmer(C_unit ~ DataSource + treatment + timestep + (1|unit), out_0506_L_C_M2)
summary(allfactors)

allfactors2 <- lmer(C_unit ~ DataSource + treatment + (1|unit), out_0506_L_C_M2)
summary(allfactors2)

allfactors3 <- lmer(C_unit ~ DataSource * treatment + (1|unit/timestep), out_0506_L_C_M2)
summary(allfactors3)

# p-value for the interaction term between data source and low residual BA is 0.04

allfactors4 <- lmer(C_unit ~ DataSource * timestep + treatment + (1|unit), out_0506_L_C_M2)
summary(allfactors4)

allfactors5 <- lmer(C_unit ~ DataSource + treatment + (1|unit/timestep), out_0506_L_C_M2)
summary(allfactors5)

allfactors6 <- lmer(C_unit ~ DataSource * treatment + timestep + (1|unit), out_0506_L_C_M2)
summary(allfactors6)



############
### DEAD ###
############

out_0506_D_C <- read.csv(paste0(path_outputs, "csv/SummitLake_C_0506_D.csv"))
summit.lk.dat_C_D <- read.csv(paste0(path_outputs, "csv/SummitLake_C_field_D.csv"))



### RSS and RMSE ###

# Compare final year of simulation C to final year of field data C - DEAD ONLY
final_C_0506_D <- subset(out_0506_D_C, timestep == 27)
final_C_0506_D <- as.data.table(final_C_0506_D)

final_field_C_D <- subset(field_C_D, timestep == 27)

final_C_0506_D <- merge(final_C_0506_D, final_field_C_D, by = c("unit", "treatment", "timestep"))
names(final_C_0506_D)[names(final_C_0506_D) == "C_unit.x"] <- "C_unit"
names(final_C_0506_D)[names(final_C_0506_D) == "C_unit.y"] <- "field_C"



rss0506 <- with(final_C_0506_D, sum((C_unit - field_C)^2))
rmse0506 <- with(final_C_0506_D, sqrt(mean((C_unit - field_C)^2)))



# Data manipulation
out_0506_D_C_M <- out_0506_D_C

out_0506_D_C_M <- out_0506_D_C_M %>%
  dplyr::select(unit, timestep, treatment, C_unit)

out_0506_D_C_M <- as.data.table(out_0506_D_C_M)

out_0506_D_C_M <- unique(out_0506_D_C_M)



out_0506_D_C_M <- subset(out_0506_D_C_M, timestep == 0 | timestep ==  2 | timestep == 5 | timestep == 17 | timestep == 27)
summit.lk.dat_C_D <- as.data.table(summit.lk.dat_C_D)

out_0506_D_C_M2 <- merge(out_0506_D_C_M, summit.lk.dat_C, by = c("unit", "treatment", "timestep"))
names(out_0506_D_C_M2)[names(out_0506_D_C_M2) == "C_unit.x"] <- "sortie"
names(out_0506_D_C_M2)[names(out_0506_D_C_M2) == "C_unit.y"] <- "field"

out_0506_D_C_M2 <- melt(out_0506_D_C_M2, id.vars = c("unit", "treatment", "timestep"), measure.vars = c("sortie", "field"))
names(out_0506_D_C_M2)[names(out_0506_D_C_M2) == "variable"] <- "DataSource"
names(out_0506_D_C_M2)[names(out_0506_D_C_M2) == "value"] <- "C_unit"


#out_0506_D_C_M2 <- rbindlist(list(out_0506_D_C_M, summit.lk.dat_C), use.names = TRUE, fill = TRUE)

out_0506_D_C_M2$unit <- as.character(out_0506_D_C_M2$unit)


# Basic linear models 
datasourcelm <- lm(C_unit ~ DataSource - 1, out_0506_D_C_M2) # forces intercept through zero, tells us the two means
summary(datasourcelm)

datasourcelm2 <- lm(C_unit ~ DataSource, out_0506_D_C_M2) # does not force the intercept through zero, tells us the difference between the two means
summary(datasourcelm2)

# gives a p-value of 0.00459 meaning there is a significant difference between the data sources

t.test(out_0506_D_C_M2[out_0506_D_C_M2$DataSource == "sortie"]$C_unit, out_0506_D_C_M2[out_0506_D_C_M2$DataSource == "field"]$C_unit)
# simple t-test, tells us the two means and if there is a difference between them (p-value)




# Linear model for each timestep WITH intercept
sourcetimelm0 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 0]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 0]$DataSource)
summary(sourcetimelm0)

sourcetimelm2 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 2]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 2]$DataSource)
summary(sourcetimelm2)

sourcetimelm5 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 5]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 5]$DataSource)
summary(sourcetimelm5)

sourcetimelm17 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 17]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 17]$DataSource)
summary(sourcetimelm17)

sourcetimelm27 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 27]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 27]$DataSource)
summary(sourcetimelm27)




# Linear model for each timestep WITHOUT intercept
sourcetimelm0 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 0]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 0]$DataSource - 1)
summary(sourcetimelm0)

sourcetimelm2 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 2]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 2]$DataSource - 1)
summary(sourcetimelm2)

sourcetimelm5 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 5]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 5]$DataSource - 1)
summary(sourcetimelm5)

sourcetimelm17 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 17]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 17]$DataSource - 1)
summary(sourcetimelm17)

sourcetimelm27 <- lm(out_0506_D_C_M2[out_0506_D_C_M2$timestep == 27]$C_unit ~ out_0506_D_C_M2[out_0506_D_C_M2$timestep == 27]$DataSource - 1)
summary(sourcetimelm27)


### MIXED EFFECTS MODELS ###

allfactors <- lmer(C_unit ~ DataSource + treatment + timestep + (1|unit), out_0506_D_C_M2)
summary(allfactors)

allfactors2 <- lmer(C_unit ~ DataSource + treatment + (1|unit), out_0506_D_C_M2)
summary(allfactors2)

allfactors3 <- lmer(C_unit ~ DataSource * treatment + (1|unit/timestep), out_0506_D_C_M2)
summary(allfactors3)

allfactors4 <- lmer(C_unit ~ DataSource * timestep + treatment + (1|unit), out_0506_D_C_M2)
summary(allfactors4)

allfactors5 <- lmer(C_unit ~ DataSource + treatment + (1|unit/timestep), out_0506_D_C_M2)
summary(allfactors5)

allfactors6 <- lmer(C_unit ~ DataSource * treatment + timestep + (1|unit), out_0506_D_C_M2)
summary(allfactors6)

# R2 - model fit of the lm ***
# covariate estimates for DataSource and Treatment
# significant difference between control and not control
# significant difference between data source



# Data manipulation keeping data at individual tree level (Probably don't need)

#out_01_D_C <- read.csv(paste0(path_outputs, "csv/SummitLake_carbon_01.csv"))

#c_vars <- c("unit", "timestep", "C_tree", "treatment")

#out_01_D_C_M3 <- out_01_D_C[c_vars]

#out_01_D_C_M3<- as.data.table(out_01_D_C_M3)
#out_01_D_C_M3[, DataSource := "sortie"]

#summit.lk.dat_C[, DataSource := "field"]

#out_01_D_C_M4 <- rbindlist(list(out_01_D_C_M3, summit.lk.dat_C), use.names = TRUE, fill = TRUE)

#out_01_D_C_M4 <- subset(out_01_D_C_M4, C_tree != "NA")


allfactors <- lmer(C_tree ~ DataSource + treatment + timestep|unit, out_01_D_C_M4)

summary(allfactors)


  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
   #     strip.background = element_blank(),strip.text.x = element_text(face="bold"), 
    #    text=element_text(size=21)) #for presentations

```



Comparison between Interior spruce and Subalpine fir
```{r}

out_0506_L_C <- read.csv(paste0(path_outputs, "csv/SummitLake_C_0506_L.csv"))

###################
### Spruce LIVE ###
###################

# Subset for Spruce
out_0506_L_C_Sx <- subset(out_0506_L_C, out_0506_L_C$Species == "Sx")

# Recalculate C per unit for spruce only
out_0506_L_C_Sx <- out_0506_L_C_Sx %>% group_by(timestep, unit) %>% mutate(C_unit = sum(C_tree, na.rm = TRUE))

# Select desired columns
out_0506_L_C_Sx <- out_0506_L_C_Sx %>%
  dplyr::select(unit, timestep, C_unit, treatment)

out_0506_L_C_Sx <- as.data.table(unique(out_0506_L_C_Sx))

# Subset for final year
final_L_C_Sx <- subset(out_0506_L_C_Sx, timestep == 27)
final_L_C_Sx <- as.data.table(final_L_C_Sx)



# See other R script to calculate field C - no spruce in unit 19
#SL_calc_field_C(raw_data = summit.lk.dat, live = TRUE, spp = "Sx")
SL_field_C_Sx <- read.csv("./Outputs/csv/SummitLake_C_field_L_Sx.csv")

# Subset for final year
final_field_C_L_Sx <- subset(SL_field_C_Sx, timestep == 27)


# Merge SORTIE data and field data 
final_L_C_Sx <- merge(final_L_C_Sx, final_field_C_L_Sx, by = c("unit", "treatment", "timestep"))
names(final_L_C_Sx)[names(final_L_C_Sx) == "C_unit.x"] <- "C_unit" # SORTIE data
names(final_L_C_Sx)[names(final_L_C_Sx) == "C_unit.y"] <- "field_C" # verify the SORTIE and field columns are being labelled correctly


lm_C_finalyrSx <-lm(C_unit ~ field_C - 1, final_L_C_Sx)
summary(lm_C_finalyrSx)

rssSx <- with(final_L_C_Sx, sum(C_unit - field_C)^2)
rmseSx <- with(final_L_C_Sx, sqrt(mean(C_unit - field_C)^2))


# Plot Expected vs. Observed for final year of carbon for spruce
png("./Figures/Final_L_C_Sx.png", width=700, height=500, units="px", pointsize=23, antialias="default")

ggplot(final_L_C_Sx, aes(x = field_C, y = C_unit, shape = treatment, color = treatment, label = unit)) +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 14), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12)) +
  geom_point(aes(colour = factor(treatment), shape = factor(treatment)), size = 2) +
  geom_text(hjust=-0.5, vjust=-0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = 1) +
  scale_x_continuous(breaks = seq(0, 60, by = 10), limits = c(0, 60)) +
  scale_y_continuous(breaks = seq(0, 60, by = 10), limits = c(0, 60)) +
  labs(x = "Observed Carbon (Mg/ha)", y = "Predicted Carbon (Mg/ha)", col = "Treatment", shape = "Treatment") +
  scale_shape_manual(values = c(16, 15, 17),
                     breaks = c("ctrl", "med", "low"), 
                     labels = c("Control", "Medium", "Low")) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium", "Low")) +
  geom_abline(slope = coef(lm_C_finalyrSx)[["field_C"]], linetype = 3)

dev.off()



##########################
### Subalpine fir LIVE ###
##########################

# Subset for Subalpine fir
out_0506_L_C_Bl <- subset(out_0506_L_C, out_0506_L_C$Species == "Bl")

# Recalculate C per unit for subalpine fir only
out_0506_L_C_Bl <- out_0506_L_C_Bl %>% group_by(timestep, unit) %>% mutate(C_unit = sum(C_tree, na.rm = TRUE))

# Select desired columns
out_0506_L_C_Bl <- out_0506_L_C_Bl %>%
  dplyr::select(unit, timestep, C_unit, treatment)

out_0506_L_C_Bl <- as.data.table(unique(out_0506_L_C_Bl))

# Subset for final year
final_L_C_Bl <- subset(out_0506_L_C_Bl, timestep == 27)

final_L_C_Bl <- as.data.table(final_L_C_Bl)


# See other R script to calculate field C
SL_calc_field_C(raw_data = summit.lk.dat, live = TRUE, spp = "Bl")
SL_field_C_Bl <- read.csv("./Outputs/csv/SummitLake_C_field_L_Bl.csv")

# Subset for final year
final_field_C_L_Bl <- subset(SL_field_C_Bl, timestep == 27)


# Merge SORTIE data and field data 
final_L_C_Bl <- merge(final_L_C_Bl, final_field_C_L_Bl, by = c("unit", "treatment", "timestep"))
names(final_L_C_Bl)[names(final_L_C_Bl) == "C_unit.x"] <- "C_unit" # SORTIE data
names(final_L_C_Bl)[names(final_L_C_Bl) == "C_unit.y"] <- "field_C" # verify the SORTIE and field columns are being labelled correctly



lm_C_finalyrBl <-lm(C_unit ~ field_C - 1, final_L_C_Bl)
summary(lm_C_finalyrBl)

rssBl <- with(final_L_C_Bl, sum(C_unit - field_C)^2)
rmseBl <- with(final_L_C_Bl, sqrt(mean(C_unit - field_C)^2))



# Plot Expected vs. Observed for final year carbon for fir
png("./Figures/Final_L_C_Bl.png", width=700, height=500, units="px", pointsize=23, antialias="default")

ggplot(final_L_C_Bl, aes(x = field_C, y = C_unit, shape = treatment, color = treatment, label = unit)) +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 14), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12)) +
  geom_point(aes(colour = factor(treatment), shape = factor(treatment)), size = 2) +
  geom_text(hjust=-0.5, vjust=-0.5) +
  geom_abline() +
  scale_x_continuous(breaks = seq(20, 90, by = 10), limits = c(20, 90)) +
  scale_y_continuous(breaks = seq(20, 90, by = 10), limits = c(20, 90)) +
  labs(x = "Observed Carbon (Mg/ha)", y = "Predicted Carbon (Mg/ha)", col = "Treatment", shape = "Treatment") +
  scale_shape_manual(values = c(16, 15, 17),
                     breaks = c("ctrl", "med", "low"), 
                     labels = c("Control", "Medium", "Low")) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "#00BA38"), 
                      breaks = c("ctrl", "med", "low"), 
                      labels = c("Control", "Medium", "Low")) +
  geom_abline(slope = coef(lm_C_finalyrBl)[["field_C"]], linetype = 3)

dev.off()


```



